---
title: "Transition from Excel Vlookup"
author: "Tam A Le"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is a guide of the R equivalent of a commonly used functions in Excel for data wrangling. I hope that this will allow people like me who have used Excel for most our carrer to transition over to R. There is nothing like getting a quick win under your belt to build your confidence.



## Excel Functions 

- Vlookup is simple,elegant and powerful function as describe by MS Excel help file;

-   In its simplest form, the VLOOKUP function says: 

-   =VLOOKUP(Value you want to look up, range where you want to lookup the     value, the column number in the range containing the return value, Exact Match or Approximate Match - indicated as 0/FALSE or 1/TRUE). 

## Problem Description

It is not often that you are handed a complete set of data, more often than not you are given fragments of data and you are expected to make sense of it and stitch them together prior to analysis. 


## Data Examples

The table below contains Total Revenue for each salesperson at each location.

```{r, echo=FALSE, results='asis'}
Revenue<-read.csv("rev.csv")
knitr::kable(head(Revenue, 9))
```


The challenge is to incorporate the above table into the table below

```{r, echo=FALSE, results='asis'}
Report<-read.csv("report.csv")
knitr::kable(head(Report, 9))
```

## R functions

Merge is an R function use to merge to two data frames together using primary key.

It is not often that you are handed a complete set of data, more often than not you are given fragments of data and you are expected to make sense of it and stitch them together prior to analysis. 


    merge (x,y, ...)

```{r, echo=FALSE, results='asis'}
Report2<-merge (Report,Revenue,"SalesPersonF")
knitr::kable(head(Report2, 13))
```

The above `merge` was not able to resolve the differences bettwen Kath Koh & Kath Pear and Peter Stjohn & Peter Thomas, `merge` advantage over vlookup is it can be match using multiple variables.

    merge (x,y, c("variable1", "variable2"))

```{r, echo=FALSE, results='asis'}
Report2<-merge (Report,Revenue,c("SalesPersonF","SalesPersonL"))
knitr::kable(head(Report2, 13))
```

Alternative to `merge` is `join` from `plyr` package, which can be use to join together two data frames, the join function is able to determine it requires to join based on `SalesPersonF` & `SalesPersonL`

    join(x, y, by = NULL, type = "left", match = "all")

```{r, echo=FALSE, results='asis'}
library (plyr)
Report3<-join (Report,Revenue,type='left',match='all')
knitr::kable(head(Report3, 13))
```



> Additional reading ([Package'plyr'](https://cran.r-project.org/web/packages/plyr/plyr.pdf))
